---
title: "Water Consumption in Saudi Arabia"
subtitle: ""
author:  <a href="https://neo-reliability.com">Ammar Alkhaldi</a>
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
library(readxl)
library(dplyr)
library(highcharter)
library(ggplot2)
library(ggimage)
library(tidyr)
```

# Introduction

`r newthought('In this brief report')`, I'll try to give an idea about the water consumption in Saudi Arabia.
The data in this report taken from the Ministry of Environment, Water and Agriculture open data^[[Ministry of Environment, Water and Agriculture](https://www.mewa.gov.sa/en/InformationCenter/OpenData/Pages/Home.aspx)]

The issue of water consumption poses a significant challenge as Saudi is a water-scarce country^[[World Resources Institute](http://www.wri.org/blog/2013/12/world%E2%80%99s-36-most-water-stressed-countries)].

## Water Consumption by Region in Million (2016 Data)

The map consumption in Million $m^3$.
We can see that Riyadh consume the majority of the produced water.


```{r, echo=FALSE}
mapdata <- get_data_from_map(download_map_data("countries/sa/sa-all"))



map <-read_excel("table8-4_2.xlsx", 
                 sheet = "Sheet2")

hc <- hcmap("countries/sa/sa-all", data = map, value = "value",
      joinBy = c("hc-a2", "code"), name = "Water Consumption in Million",
      dataLabels = list(enabled = TRUE, format = '{point.name}'),
      borderColor = "#FAFAFA", borderWidth = 0.1,
      tooltip = list(valueDecimals = 2, valuePrefix = "", valueSuffix = "m³"))

hc_colorAxis(hc, minColor = "#f3eded", maxColor = "#aa530e")

```

While the map show high consumption in Riyadh we still not sure if that consumbtion was due to the high population or not, to have a deeper understanding the water consumbution will be done per capita.

# Consumption per Captia per Day {.no-padding}

`r newthought('By looking to the figure below')`, We can see the variance in the water consumpition between different regions, for example the average water usage in eastren region is 4X times the water consumpition in Asir.





```{r echo=FALSE, fig.height=12, fig.width=12, warning=FALSE, dev='Cairo_svg'}

region <- c("Ar Riyad","Makkah", "Madinah", "Al Quassim", "Ash Sharqiyah", "Asir",
            "Tabuk", "Ha'il", "Al Hudud ash Shamaliyah", "Jizan", "Najran",
            "Al Bahah", "Al Jawaf")
ratio <- readRDS("ratio.rds")
df1 <- data.frame(grp=as.character(region),
                  img=rep("./picto.png",13),
                  freq=as.numeric(ratio))
df1 <- arrange(df1,freq)
df1$y <- 13:1

df3 <- df1[rep(seq_len(nrow(df1)), df1$freq), c(1,2,4)]
df3 <- df3 %>% group_by(grp) %>% mutate(x = row_number())
df3$x <- df3$x + 1

ggplot(df3) +
  geom_image(aes(x=x, y=y, image=img),size=.06) +
  theme_void() +
  geom_text(data=df1, aes(y=y+0.2, label=freq), x=-0.1, 
            fontface=2, size=10, color="#F26821") +
  geom_text(data=df1, aes(y=y-.25, label=grp), x=-0.1, 
            fontface=2, size= 8,family= "Garamond") +
  xlim(-1,max(df1$freq)+2) + ylim(0,max(df1$y)+1)+
  geom_rect(aes(xmin = 2.5, xmax = 6, ymin = 11.5, ymax = 13.5),
            fill = "transparent", color = "black", size = 1.5) +
  geom_image(aes(x=3,y=12.5, image=img))+
  geom_text(aes(x=4.6, y=12.5,
                label="= 108 (baseline)"),
            size= 10,
            color="#636363",
            family= "Garamond")+
  geom_hline(yintercept = seq(1.5,11.5,1),linetype = 5, size= 0.5)+
  geom_segment(x=-2,xend=2.5,y=12.5,yend=12.5,linetype = 5, size= 0.5)+
  labs(title = "Daily Water Consumpition Per Capita (L/day)")+
  theme(plot.background = element_rect(fill = "#fffff8"),
        text=element_text(size=21 ,family="Garamond"))

```



```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
